Postural changes mediate children's visual access to social information

### 2018-01-15 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
renaming directories

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and then
using modern computer-vision algorithms to detect the proportion of
faces and hands in infants' environments. We find that infant's postural
changes largely mediate infants' acess to both faces and hands,
suggesting that these postural and locomotive developments facilitate
infants' emerging linguistic and social capacities. Broadly, we suggest
that the combined use of head-mounted cameras and the application of
novel deep learning algorithms is a promising avenue for understanding
the statistics of infants' visual and linguistic experience.

@@

Here, we directly ask whether the postural changes that infants'
experience as they reach their first birthday change the availability of
the social information relevant for word learning. To do so, we recorded
the visual experience of a group of infants and children using
head-mounted cameras during a brief laboratory free-play session;
children's posture and their orientation to their caregiver were also
recorded from a third-person perspective and hand-annotated. We
capitalize on recent improvements in computer vision algorithms (K.
Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) that
allow the automated detection of both faces and hands, and we analyze
the frequency of both faces and hands in the child's visual environment.

@@

The second of these algorithms (OpenPose) also captures information
about the ``skeleton'' of people in a given frame, outputting
information about the entire set of body parts that were detected. This
allowed us to extract hand information as well as wrist information. We
analyze wrist detections as a good proxy for the presence of a hand. We
do so first because we did not capture the entire visual field
experienced by the infants/children. Thus evne though the video may only
show the presence of a wrist, the caregiver's hand may be within the
child's field of view. Second, we do so because hands are often occluded
by objects when caregivers are interacting with children, yet still
visually accessible by the child and part of their joint interaction.
For example,if a caregiver was holding a toy and presenting it to the
infant, their wrists may be visible but their hand may not necessarily
be (as it is occluded by the toy).

@@

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc) (Cao et al. (2017),
Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna, Kanade, \&
Sheikh (2016)) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose). We operationalized face detections as any
frames containing a nose and either the left or right eye; we
operationalized hand detections as any frames containing any of the 21
points detected on either the left or the right hand, and we
operationalized wrist detections any frames containing either the left
or the right wrist.

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.84 & 1.00 & 0.91 \\ 
   \hline
\end{tabular}
\caption{MTCNN High Density Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.95 & 0.75 & 0.84 \\ 
   \hline
\end{tabular}
\caption{MTCNN Random Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.99 & 0.87 & 0.93 \\ 
   \hline
\end{tabular}
\caption{OpenPose High Density Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.71 & 0.92 & 0.80 \\ 
   \hline
\end{tabular}
\caption{OpenPose Random Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.65 & 0.85 & 0.74 \\ 
   \hline
\end{tabular}
\caption{OpenPose Wrist High Density Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.32 & 0.13 & 0.19 \\ 
   \hline
\end{tabular}
\caption{OpenPose Wrist Random Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.94 & 0.65 & 0.77 \\ 
   \hline
\end{tabular}
\caption{OpenPose Hand High Density Sample} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & P & R & F \\ 
  \hline
1 & 0.86 & 0.13 & 0.23 \\ 
   \hline
\end{tabular}
\caption{OpenPose Hand Random Sample} 
\end{table}

Evaluating hand/wrist detectors (text here about how gold standard was
selected)

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{TABLE PLACEHOLDER} 
\end{table}

@@

First, we report developmental shifts in infants' posture and their
orientation to their caregiver. Then, we explore how these changes
influence children's visual access to faces and to hands. Finally, we
explore how these changes impact the accessibility of faces and hands
during labeling events.

@@

\section{Changes in Access to Faces \&
Hands}\label{changes-in-access-to-faces-hands}

We first examined the proportion of face and hand detections across age;
a full summary can be seen in Figure \ref{fig:detsbyAge}. We observed a
slight u-shaped function both when analyzing the output of the MTCNN and
OpenPose detectors, such that 12-month-olds may experience less faces
than 8 or 12-month-olds. Next, we observed that hand detections
increased with the age of the participant, though they were relatively
infrequent in this dataset. As expected, we found that wrist detections
we relatively more frequent and also appeared to increase somewhat with
age, suggesting that caregivers hands may be just outside of the field
of view of our camera or that they were often occluded by toys.

Next, we turned to our hypothesis of interest, examinig the impact of
postural and locomotive developments on children's visual access to
faces and hands. Overall, we ofund that children's posture was a major
factor both in how many faces they saw during the play session. Infants
who were sitting saw more faces than infants who were lying down or
being carried, while infants who were standing saw the most faces. We
observed a similar pattern with respect to hands: children who were
sitting or standing tended to see more hands/wrists than children who
were lying down or held.

We next examined how the child's orientation relative to their caregiver
impacted their visual access to faces and hands. Overlal, we saw that
children who were far away from their caregiver were more likely to see
faces and hands than children who were close to their caregiver; this
was true within all age groups and for all detectors.

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -8.775 & 0.236 & -37.259 & 0.000 \\ 
  Age & 0.879 & 0.166 & 5.292 & 0.000 \\ 
  Posture-Prone  & 1.215 & 0.110 & 11.082 & 0.000 \\ 
  Posture-Sit & 1.943 & 0.108 & 18.000 & 0.000 \\ 
  Posture-Stand  & 2.175 & 0.107 & 20.260 & 0.000 \\ 
  Orient.-Close  & 0.608 & 0.059 & 10.371 & 0.000 \\ 
  Orient.-Far & 2.244 & 0.060 & 37.222 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.} 
\end{table}

@@

We use a head-mounted camera to explore how children's postural and
locomotive development directly impacts their access to social
information relevant for word-learning, here operationalized as the
presence of the faces and hands of their caregiver. We found that
children's posture and orientation towards their caregiver changed
systematically across age, and that all three of these factors
dramatically impacted the proportion of faces and hands that were
avaliable in the child's visual field. Thus, infants postural
development is mediating factor that explains age-related changes in the
proportion of faces and hands avaliable in infants' visual field.

This work also deploys novel advancements in computer vision to the
study of developmental psychology. The field of object detection and
recognition has advanced dramatically in the past five years since the
re-birth of deep learning algorithms Krizhevsky, Sutskever, \& Hinton
(2012), creating a new generation of algorithmic tools. These tools are
both substantially better equipped to deal with noisier, more
complicated datasets and that can extract richer and more detailed
information. Videos from the infant perspective provided sustantial
challenges (e.g., partially occluded faces) for the classc models of
face detedction (e.g., ViolaJones, Viola \& Jones (2004)). urther, as
the headcam technologies employed here were inexpensive
(\textasciitilde{}\$60 a camera) and the computer vision algorithms
freely avaliable, this method is a promising avenue for quanitfying the
visual and social information avaliable to infant learners.

>>>>

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and then
using modern computer-vision algorithms to detect the proportion of
faces in infants' environments. We find that infant's postural changes
largely mediate infants' acess to faces, suggesting that these postural
and locomotive developments facilitate infants' emerging linguistic and
social capacities. Broadly, we suggest that the combined use of
head-mounted cameras and the application of novel deep learning
algorithms is a promising avenue for understanding the statistics of
infants' visual and linguistic experience.

@@

Here, we directly ask whether the postural changes that infants'
experience as they reach their first birthday change the availability of
the social information relevant for word learning. To do so, we recorded
the visual experience of a group of infants and children using
head-mounted cameras during a brief laboratory free-play session;
children's posture and their orientation to their caregiver were also
recorded from a third-person perspective and hand-annotated. We
capitalize on recent improvements in computer vision algorithms (K.
Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) that
allow the automated detection of both faces and hands, and we analyze
the frequency of faces in the child's visual environment.

@@

@@

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc) (Cao et al. (2017),
Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna, Kanade, \&
Sheikh (2016)) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose). We operationalized face detections as any
frames containing a nose, as any half of a face necessarily contains a
nose.\\

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
  & P & R & F \\ 
\hline
OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 
OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\
MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\ 
MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\ 
   \hline
\end{tabular}
\caption{Face detector performance} 
\end{table}

@@

First, we report developmental shifts in infants' posture and their
orientation to their caregiver. Then, we explore how these changes
influence children's visual access to faces across this developmental
time range. Finally, we explore how these changes impact the
accessibility of faces during labeling events.

@@

\section{Changes in Access to Faces}\label{changes-in-access-to-faces}

We first examined the proportion of face detections across age; a full
summary can be seen in Figure \ref{fig:detsbyAge}. We observed a slight
u-shaped function both when analyzing the output of the MTCNN and
OpenPose detectors, such that 12-month-olds appeared to experience
slighly fewer faces faces than 8 or 16-month-olds.

Next, we turned to our hypothesis of interest, examinig the impact of
postural and locomotive developments on children's visual access to
faces. Overall, we ofund that children's posture was a major factor both
in how many faces they saw during the play session. Infants who were
sitting saw more faces than infants who were lying down or being
carried, while infants who were standing saw the most faces. We next
examined how the child's orientation relative to their caregiver
impacted their visual access to faces. Overall, we saw that children who
were far away from their caregiver were more likely to see faces than
children who were close to their caregiver; this was true within all age
groups and for all detectors.

@@

\subsection{Access to Faces During Labeling
Events}\label{access-to-faces-during-labeling-events}

Finally, we analyzed how face detections changed during object labeling
events as a function of infant's posture and orientation. Specfically,
we analyzed a four-second window around each labeling event (e.g.,
``Look at the {[}zem{]}!''); these labelling events were hand-annotated
and automatically synchronized with the frame-by-frame face detections.
We again found that infant's posture the degree to which they saw their
caregivers face during a labeling event; infants who were sitting or
standing were more likely to have access to their caregivers face.

{\centering \includegraphics{figs/detByNaming-1}

\caption[Proportion face detections as a function of infants's posture]{Proportion face detections as a function of infants's posture}\label{fig:detByNaming}
\end{figure}
\end{CodeChunk}

@@

We use a head-mounted camera to explore how children's postural and
locomotive development directly impacts their access to social
information relevant for word-learning, here operationalized as the
presence of the faces of their caregiver. We found that children's
posture and orientation towards their caregiver changed systematically
across age, and that all three of these factors dramatically impacted
the proportion of faces that were avaliable in the child's visual field.
Thus, infants postural development is mediating factor that explains
age-related changes in the proportion of faces avaliable in infants'
visual field.

This work also deploys novel advancements in computer vision to the
study of developmental psychology. The field of object detection and
recognition has advanced dramatically in the past five years since the
re-birth of deep learning algorithms Krizhevsky, Sutskever, \& Hinton
(2012), creating a new generation of algorithmic tools. These tools are
both substantially better equipped to deal with noisier, more
complicated datasets and that can extract richer and more detailed
information. Videos from the infant perspective provided sustantial
challenges (e.g., partially occluded faces) for the classc models of
face detedction (e.g., ViolaJones, Viola \& Jones (2004)). Further, as
the headcam technologies employed here were inexpensive
(\textasciitilde{}\$60 a camera) and the computer vision algorithms
freely avaliable, this method is a promising avenue for quanitfying the
visual and social information avaliable to infant learners.
